---
title: "Project 2"
author: "LeTicia Cancel"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

```

# Dataset #1
##### Source: [New York State Education Department](https://data.nysed.gov/downloads.php)
##### Data Description: ELA and Math test scores for grades 3-8 by state and county
**Note:** The original file is 57MB so I did minor manipulation using excel to create a smaller file. I copy and pasted all columns for the statewide data and each NYC county data to a new csv file. This is the only data manipulation done outside of R and this subset of data will be used in this project.

--------
#### Load all libraries
```{r warning=FALSE, results='hide', message=FALSE}
#install.packages("kableExtra") #install this package if it is not already installed
library(knitr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RCurl)
library(kableExtra)
```


Load csv file
```{r}
df <- read.table(text = getURL("https://raw.githubusercontent.com/ltcancel/Project2/master/3-8%20ELA%20and%20Math%20Scores%20by%20County.csv"), header = TRUE, sep = ",")
```



# Dataset #2
##### Source: [Open Intro](https://openintro.org/stat/data/pokemon.csv)
##### Data Description:
**Note:** A csv file containing Pokemon data will download to your machine when you click on the Open Intro link



# Dataset #3
##### Source: [UFC-Fight historical data from 1993 to 2019](https://www.kaggle.com/rajeevw/ufcdata/version/2#raw_total_fight_data.csv)
##### Data Description: 