---
title: "Project 2"
author: "LeTicia Cancel"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dataset #1
##### Source: [New York State Education Department](https://data.nysed.gov/downloads.php)
##### Data Description: ELA and Math test scores for grades 3-8 by state and county
**Note:** The original file is 57MB so I did minor manipulation using excel to create a smaller file. I copy and pasted all columns for the statewide data and each NYC county data to a new csv file. This is the only data manipulation done outside of R and this subset of data will be used in this project.

--------
#### Load all libraries
```{r warning=FALSE, results='hide', message=FALSE}
#install.packages("kableExtra") #install this package if it is not already installed
library(knitr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RCurl)
library(kableExtra)
library(tidyverse)
```


Load csv file
```{r}
df <- read.table(text = getURL("https://raw.githubusercontent.com/ltcancel/Project2/master/3-8%20ELA%20and%20Math%20Scores%20by%20County.csv"), header = TRUE, sep = ",")
head(df)
```

#### Create dataframes based on subgroups.

Gender subgroup
```{r}
gender.df <- df %>% filter(SUBGROUP_CODE %in% c(1:3))
head(gender.df)
```

Race/Ethnicity subgroup
```{r}
race.df <- df %>% filter(SUBGROUP_CODE == "1" | SUBGROUP_CODE %in% c(4:9))
head(race.df)
```

Education subgroup
```{r}
edu.df <- df %>% filter(SUBGROUP_CODE == "1" | SUBGROUP_CODE %in% c(10:11))
head(edu.df)
```

Language subgroup
```{r}
language.df <- df %>% filter(SUBGROUP_CODE == "1" | SUBGROUP_CODE %in% c(12:13))
head(language.df)
```

Economic subgroup
```{r}
eco.df <- df %>% filter(SUBGROUP_CODE =="1" | SUBGROUP_CODE %in% c(15:16))
head(eco.df)
```


# Dataset #2
##### Source: [Open Intro](https://openintro.org/stat/data/pokemon.csv)
##### Data Description:
**Note:** A csv file containing Pokemon data will download to your machine when you click on the Open Intro link

Load csv file
```{r}
df2 <- read.csv(text = getURL("https://raw.githubusercontent.com/ltcancel/Project2/master/pokemon.csv"), header = TRUE, sep = ",")
df2
?read.csv
```




# Dataset #3
##### Source: [UFC-Fight historical data from 1993 to 2019](https://www.kaggle.com/rajeevw/ufcdata/version/2#raw_total_fight_data.csv)
##### Data Description: 

Load csv file
```{r}
df3 <- read.csv(text = getURL("https://raw.githubusercontent.com/ltcancel/Project2/master/raw_total_fight_data.csv"), header = TRUE, sep = ";")
head(df3)
```

